---
title: "Student project- Data exploration"
author: "Luke Henslee"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

### Import data
```{r echo=TRUE, results=FALSE}

library(tidyverse)

setwd("C:/Users/lhhenslee/Desktop/Luke/School/Coursework/604/data")

# Import mcode list

  # The mcode list contains capture data (including environmental observations), recapture information, acoustic detection history, and final fates for each tagged salmon

    # 'colClasses.csv' is just a list for 'colClasses' argument of read.csv() 
colClasses <- read.csv("colClasses.csv", header = T)

    # Import data
mcode <- read.csv("mcode.csv", colClasses = paste(colClasses[1, ])) 


# Import cpue dataset

  # This dataset contains location, environmental observations, and CPUE for each sample attempt
cpue <- read.csv("cpue.csv")

```


### Data exploration

#### mcode data
```{r echo=TRUE}
# Let's look at the mcode data

str(mcode)
head(mcode)

```

Could perhaps use some data reduction.... 

```{r}
# Data visualization

ggplot(data = mcode, aes(x = spawn.group)) +
  geom_bar()

ggplot(data = mcode, aes(x = final.fate)) +
  geom_bar(aes(fill = tag.group))

ggplot(data = mcode, aes(x = capture.date)) +
  geom_bar(aes(fill = tag.group))

```

#### CPUE data
```{r echo=TRUE}
# Let's look at the cpue data

str(cpue)
head(cpue)

```

```{r}
# Data visualization

plot(cpue$julian.date, cpue$CPUE)

ggplot(data = cpue, aes(julian.date, CPUE)) +
  geom_point() +
  geom_smooth()

ggplot(data = cpue, aes(julian.date, CPUE)) +
  geom_point(aes(color = site)) +
  geom_smooth(aes(color = site), method = "loess")


```

### Playing around with GLM
```{r}

ggplot(data = cpue, aes(x = lat, y = fish)) +
  geom_point() +
  stat_smooth(method = 'glm',  method.args = list(family = "poisson"),
              formula = y ~ x + I(x^2))

# Minus the outlier
outlier <- cpue %>% 
  filter(lat < 65.5)

ggplot(data = outlier, aes(x = lat, y = fish)) +
  geom_point() +
  stat_smooth(method = 'glm',  method.args = list(family = "poisson"),
              formula = y ~ x + I(x^2))

```


### Question 1- Can we identify a spatiotemporal covariate to estimate probibility of stock membership?

My idea is to use CPUE metrics to estimate stock-specific catch rate parameters in different spatial and temporal strata. It might be possible to use a non-linear model with a Poisson distribution to describe catch rates.

Are there autocorrelation issues based on distance between capture location and homestream? Or between week of capture and peak of stock run?

CPUE is not count data, does Poisson work?

### Question 2- Can we estimate stock proportions of commercial harvest?

Based on the results from question 1, can we apply stock-specific catch rates to commercial harvests in spatial and temporal strata to estimate catch stock proportions?

The goal is to evaluate weather stock-specific harvest strategies can be formulated based on probabilities of stock-specific capture in several spatial and temporal strata (i.e. commercial fishing subdistricts, ADF&G statistical weeks)

