---
title: "Student project- Data exploration"
author: "Luke Henslee"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r, include=FALSE, results=FALSE}

library(ggpubr)
library(forcats)
library(ggforce)
library(lubridate)
library(RColorBrewer)
library(viridis)
library(tidyverse)
library(ggplot2)
library(bbmle)
library(dplyr)
library(manipulate)
library(visreg) 
library(ggridges)
library(reshape2)
library(cowplot)
library(ggmap)
library(ggspatial)
library(sf)
library(spData)
library(ggsn)
library(extrafont)
loadfonts(device = "win")

```

```{r include=FALSE}
setwd("C:/Users/lukeh/Desktop/Grad School/Coursework/604/HW/HW2")

mcode <- read.csv("sablefish.csv", as.is=F)

library(ggplot2)
library(tidyverse)
```

### Question 1- Can we identify a spatiotemporal covariate to estimate probibility of stock membership?

My idea is to use CPUE metrics to estimate stock-specific catch rate parameters in different spatial and temporal strata. It might be possible to use a non-linear model with a Poisson distribution to describe catch rates

### Question 2- Can we estimate stock proportions of commercial harvest?

