---
title: "Final Assignment"
author: "Luke Henslee"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r, include=FALSE, results=FALSE}
library(tidyverse)
library(ggplot2)
library(extrafont)
library(ggpubr)
library(forcats)
library(ggforce)
library(ggridges)
library(lubridate)
library(RColorBrewer)
library(viridis)
library(readr)
# Load fonts
loadfonts(device = "win")
```

### Import data   

```{r, root.dir}
setwd("C:/Users/lukeh/Desktop/Grad School/Coursework/635/data")

mcode <- read.csv("M_code_list.csv")

# Subset data to fish assigned to stocks
stocks <- filter(mcode, mcode$Spawning.group %in% c("4", "5", "6", "S"))

# Convert 'Date'
stocks$Date <- mdy(stocks$Date)
```

### Figure 1- Length distribution

#### 1) Build plot  

First we plot length distribution between the four identified stocks.
```{r}
# Build 'lengthdist'
lengthdist <- ggplot(data = stocks, aes(x = Spawning.group, y = Length)) +
geom_boxplot() +
  scale_x_discrete(breaks = c("6", "5", "4", "S"), 
    labels = c("Unalakleet", "Shaktoolik", "Norton Bay", "Yukon")) +
scale_y_continuous (limits = c(375,700), expand = c(0,0), breaks = seq(400,700,50)) +
ylab("Length (mm)") +
theme_classic() +
theme(axis.title.y = element_text(size = 14, margin = margin(t = 0, r = 10, b = 0, l = 0), colour = "black"),
      axis.title.x = element_blank(),
      text = element_text(family = "Times New Roman"),
      plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),
      axis.ticks.y = element_line(size = 0.5),
      axis.ticks.x = element_line(size = 0.5),
      axis.ticks.length = unit(0.2,"cm"),
      axis.text.y = element_text(colour = "black", size = 12, angle = 0, vjust = 0.5, hjust = 1,
          margin = margin(t = 0, r = 5, b = 0, l = 0)),
      axis.text.x = element_text(colour = "black", size = 14, angle = 0, vjust = 0, hjust = 0.5,
          margin = margin(t = 5, r = 0, b = 0, l = 0)),
      axis.line = element_line(colour = "black", size = 0.5, lineend = "square"))
lengthdist
```

Save plot

```{r}
dev.new(width = 2.75, height = 1.83, units = "in", noRStudioGD = TRUE)
last_plot()

ggsave(file = "Final_lengthdist_1.tiff", width = dev.size()[1], height = dev.size()[2], dpi = 300)
```

Let's try a length distribution plot using just stocks 5, 6, and S, since there were only four samples for stock 4.

```{r}
stocks2 <- filter(mcode, mcode$Spawning.group %in% c("5", "6", "S"))

lengthdist2 <- ggplot(data = stocks2, aes(x = Spawning.group, y = Length)) +
geom_boxplot() +
  scale_x_discrete(breaks = c("6", "5", "S"), 
    labels = c("Unalakleet", "Shaktoolik", "Yukon")) +
scale_y_continuous (limits = c(375,700), expand = c(0,0), breaks = seq(400,700,50)) +
ylab("Length (mm)") +
xlab("Stocks") +
theme_classic() +
theme(axis.title.y = element_text(size = 14, margin = margin(t = 0, r = 10, b = 0, l = 0), colour = "black"),
      axis.title.x = element_blank(),
      text = element_text(family = "Times New Roman"),
      plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),
      axis.ticks.y = element_line(size = 0.5),
      axis.ticks.x = element_line(size = 0.5),
      axis.ticks.length = unit(0.2,"cm"),
      axis.text.y = element_text(colour = "black", size = 12, angle = 0, vjust = 0.5, hjust = 1,
          margin = margin(t = 0, r = 5, b = 0, l = 0)),
      axis.text.x = element_text(colour = "black", size = 12, angle = 0, vjust = 0, hjust = 0.5,
          margin = margin(t = 5, r = 0, b = 0, l = 0)),
      axis.line = element_line(colour = "black", size = 0.5, lineend = "square"))
lengthdist2
```

Really not that interesting, I don't think I'm going to use it.

### Figure 2- Catch proportions by stock

We're going to visualize proportions of total catch by stock throughout the season.

We are using the "count" stat, is it the same as proportion/density? No, because then the whole background would be filled since it equals 100%, unless that was the daily proportion of the total. Might try to mess around witht that.

```{r}
ggplot(stocks, aes(x = yday(Date), y = ..count..)) +
  geom_density_line(data = select(stocks, -Spawning.group), color = "transparent") +
  geom_density_line(fill = "blue", alpha = 0.5, color = "transparent") +
  facet_wrap(~Spawning.group, nrow = 4) +
  scale_x_continuous(name = "Day of year", limits = c(213,239), expand = c(0,0), breaks = seq(215,235,5))
```

I'm going to save this one.
```{r}
dev.new(width = 2.75, height = 1.83, units = "in", noRStudioGD = TRUE)
last_plot()

ggsave(file = "Final_stock_count_wrap.tiff", width = dev.size()[1], height = dev.size()[2], dpi = 300)
```

Try to visualize all in one plot.

```{r}
ggplot(stocks, aes(x = yday(Date), y = ..count..)) +
  geom_density_line(data = select(stocks, -Spawning.group), alpha = 0.5, color = "darkgrey", fill = "lightgrey", size = 0.25) +
  geom_density_line(aes(fill = fct_infreq(Spawning.group)), alpha = 0.5, color = "black", size = 0.25) +
  scale_x_continuous(name = "Day of year", limits = c(213,239), expand = c(0,0), breaks = seq(215,235,5)) +
  scale_y_continuous(name = "Count", expand = c(0, 0), limits = c(0, 22), breaks = seq(5, 20, 5)) +
  scale_fill_brewer(type = "qual", guide = "legend",
    palette = "Set2", name = "Stocks", breaks = c("6", "5", "4", "S"), 
    labels = c("Unalakleet", "Shaktoolik", "Norton Bay", "Yukon")) +
  theme_classic() +
theme (axis.title.y = element_text(size = 14, margin = margin(t = 0, r = 10, b = 0, l = 0), colour = "black"),
      axis.title.x = element_text(size = 14, margin = margin(t = 10, r = 0, b = 0, l = 0), colour = "black"),
      text = element_text(family = "Times New Roman"),
      legend.position = c(0.75,0.75),
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 12),
      legend.key.size = unit(.5, "cm"),
      plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
      axis.ticks.y = element_line(size = 0.5),
      axis.ticks.x = element_line(size = 0.5),
      axis.ticks.length = unit(0.2,"cm"),
      axis.text.y = element_text(colour = "black", size = 14, angle = 0, vjust = 0.5, hjust = 1,
          margin = margin(t = 0, r = 5, b = 0, l = 0)),
      axis.text.x = element_text(colour = "black", size = 14, angle = 0, vjust = 0, hjust = 0.5,
          margin = margin(t = 5, r = 0, b = 0, l = 0)),
      axis.line = element_line(colour = "black", size = 0.5, lineend = "square"),
      panel.border = element_rect(color = "black", fill = "transparent", size = 0.5))
```
```{r}
dev.new(width = 20.32, height = 7.62, units = "cm", noRStudioGD = TRUE)
last_plot()

ggsave(file = "Final_density_V5.tiff", width = dev.size()[1], height = dev.size()[2], dpi = 300)
```

# Figure 2 

```{r}
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)
library(sf)
library(rgeos)
library(ggmap)

world <- ne_countries(scale = "medium", returnclass = "sf")
ggplot(data = world) +
        theme_classic() +  
        geom_sf() +
        coord_sf(xlim = c(-161.5, -160.5), ylim = c(63.5, 64.5), expand = TRUE)  +
        geom_point(data=stocks,
                   aes(x=Long, y=Lat, color=Spawning.group), alpha=0.5) +
        scale_color_viridis(discrete = TRUE, option = "C") 
```

```{r}
SD5 <- get_map(location=c(-161.6, 64.1,
                              -160.75, 64.5),
                   maptype='toner-lite', source='stamen', crop=TRUE)

SD56 <- get_map(location=c(-161.6, 63.8,
                              -160.75, 64.5),
                   maptype='toner-lite', source='stamen', crop=TRUE)
```


```{r}
ggmap(SD5) +
        geom_point(data = stocks5, aes(x=Long, 
                       y=Lat, color = Spawning.group), alpha = 0.5, size = 3) +
  scale_fill_brewer(
    palette = "Set2", name = "Stocks", breaks = c("6", "5", "4", "S"), 
    labels = c("Unalakleet", "Shaktoolik", "Norton Bay", "Yukon")) +
  theme_classic() +
theme (axis.title.y = element_text(size = 14, margin = margin(t = 0, r = 10, b = 0, l = 0), colour = "black"),
      axis.title.x = element_text(size = 14, margin = margin(t = 10, r = 0, b = 0, l = 0), colour = "black"),
      text = element_text(family = "Times New Roman"),
      legend.position = c(0.1,0.1),
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 12),
      plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
      axis.ticks.y = element_line(size = 0.5),
      axis.ticks.x = element_line(size = 0.5),
      axis.ticks.length = unit(0.2,"cm"),
      axis.text.y = element_text(colour = "black", size = 14, angle = 0, vjust = 0.5, hjust = 1,
          margin = margin(t = 0, r = 5, b = 0, l = 0)),
      axis.text.x = element_text(colour = "black", size = 14, angle = 0, vjust = 0, hjust = 0.5,
          margin = margin(t = 5, r = 0, b = 0, l = 0)),
      axis.line = element_line(colour = "black", size = 0.5, lineend = "square"),
      panel.border = element_rect(color = "black", fill = "transparent", size = 0.5))
```

```{r}
ggmap(SD56) +
        geom_violin(data=stocks,
                   aes(x=Spawning.group, y=Lat, color="blue")) 
```

```{r}
ggplot(data = stocks, aes(x = Spawning.group, y = Lat)) +
  geom_violin(aes(fill = fct_infreq(Spawning.group)), alpha = 0.5, color = "black", size = 0.5) +
  geom_hline(yintercept = 64.131287, linetype = "dashed") +
  scale_x_discrete(name = "", breaks = c("6", "5", "4", "S"), 
    labels = c("Unalakleet", "Shaktoolik", "Norton Bay", "Yukon")) + 
  scale_y_continuous(name = "Latitude of capture", expand = c(0, 0), limits = c(63.8, 64.4), breaks = seq(63.8, 64.4, 0.2)) +
  scale_fill_brewer(type = "qual", guide = "legend",
    palette = "Set2", name = "Stocks", breaks = c("6", "5", "4", "S"), 
    labels = c("Unalakleet", "Shaktoolik", "Norton Bay", "Yukon")) +
  theme_classic() +
theme (axis.title.y = element_text(size = 14, margin = margin(t = 0, r = 10, b = 0, l = 0), colour = "black"),
      axis.title.x = element_blank(),
      text = element_text(family = "Times New Roman"),
      legend.position = "none",
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 12),
      plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),
      axis.ticks.y = element_line(size = 0.5),
      axis.ticks.x = element_line(size = 0.5),
      axis.ticks.length = unit(0.2,"cm"),
      axis.text.y = element_text(colour = "black", size = 12, angle = 0, vjust = 0.5, hjust = 1,
          margin = margin(t = 0, r = 5, b = 0, l = 0)),
      axis.text.x = element_text(colour = "black", size = 14, angle = 0, vjust = 0, hjust = 0.5,
          margin = margin(t = 5, r = 0, b = 0, l = 0)),
      axis.line = element_line(colour = "black", size = 0.5, lineend = "square"),
      panel.border = element_rect(color = "black", fill = "transparent", size = 0.5))
```

```{r}
dev.new(width = 2.75, height = 1.83, units = "in", noRStudioGD = TRUE)
last_plot()

ggsave(file = "Final_violin.tiff", width = dev.size()[1], height = dev.size()[2], dpi = 300)
```



```{r}
ggplot(data = stocks, aes(x = Spawning.group, y = Lat)) +
  geom_violin() +
  geom_sina() +
  scale_x_discrete(breaks = c("6", "5", "4", "S"), 
    labels = c("Unalakleet", "Shaktoolik", "Norton Bay", "Yukon")) + 
  scale_y_continuous(name = "Latitude of capture", expand = c(0, 0), limits = c(63.8, 64.4), breaks = seq(63.8, 64.4, 0.2)) +
  xlab(label = "") +
  geom_hline(yintercept = 64.131287) +
  theme_classic() +
theme (axis.title.y = element_text(size = 14, margin = margin(t = 0, r = 10, b = 0, l = 0), colour = "black"),
      axis.title.x = element_text(size = 14, margin = margin(t = 10, r = 0, b = 0, l = 0), colour = "black"),
      text = element_text(family = "Times New Roman"),
      plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),
      axis.ticks.y = element_line(size = 0.5),
      axis.ticks.x = element_line(size = 0.5),
      axis.ticks.length = unit(0.2,"cm"),
      axis.text.y = element_text(colour = "black", size = 12, angle = 0, vjust = 0.5, hjust = 1,
          margin = margin(t = 0, r = 5, b = 0, l = 0)),
      axis.text.x = element_text(colour = "black", size = 12, angle = 0, vjust = 0, hjust = 0.5,
          margin = margin(t = 5, r = 0, b = 0, l = 0)),
      axis.line = element_line(colour = "black", size = 0.5, lineend = "square"),
      panel.border = element_rect(color = "black", fill = "transparent", size = 0.5))
```


```{r}
ggplot(data = stocks, aes(x = Spawning.group, y = Lat)) +
  geom_violin(aes(fill = fct_infreq(Spawning.group)), alpha = 0.5, color = "black", size = 0.5) +
  geom_hline(yintercept = 64.131287, linetype = "dashed") +
  geom_sina() +
  scale_x_discrete(name = "", breaks = c("6", "5", "4", "S"), 
    labels = c("Unalakleet", "Shaktoolik", "Norton Bay", "Yukon")) + 
  scale_y_continuous(name = "Latitude of capture", expand = c(0, 0), limits = c(63.8, 64.4), breaks = seq(63.8, 64.4, 0.2)) +
  scale_fill_brewer(type = "qual", guide = "legend",
    palette = "Set2", name = "Stocks", breaks = c("6", "5", "4", "S"), 
    labels = c("Unalakleet", "Shaktoolik", "Norton Bay", "Yukon")) +
  theme_classic() +
theme (axis.title.y = element_text(size = 14, margin = margin(t = 0, r = 10, b = 0, l = 0), colour = "black"),
      axis.title.x = element_blank(),
      text = element_text(family = "Times New Roman"),
      legend.position = "none",
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 12),
      plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),
      axis.ticks.y = element_line(size = 0.5),
      axis.ticks.x = element_line(size = 0.5),
      axis.ticks.length = unit(0.2,"cm"),
      axis.text.y = element_text(colour = "black", size = 12, angle = 0, vjust = 0.5, hjust = 1,
          margin = margin(t = 0, r = 5, b = 0, l = 0)),
      axis.text.x = element_text(colour = "black", size = 14, angle = 0, vjust = 0, hjust = 0.5,
          margin = margin(t = 5, r = 0, b = 0, l = 0)),
      axis.line = element_line(colour = "black", size = 0.5, lineend = "square"),
      panel.border = element_rect(color = "black", fill = "transparent", size = 0.5))
```

```{r}
dev.new(width = 2.75, height = 1.83, units = "in", noRStudioGD = TRUE)
last_plot()

ggsave(file = "Final_violin_sina.tiff", width = dev.size()[1], height = dev.size()[2], dpi = 300)
```

# GLM

```{r}
stocks5 <- filter(stocks, stocks$Tag.group == 2)

bi <- vector(length = nrow(stocks5))

for(i in 1:length(bi)) {
  if(stocks5[i,9] == stocks5[i,16]) {
    bi[i] <- 1
  } else {
    bi[i] <- 0
  }
}

stocks5$binomial <- bi
```


```{r}
ggplot(stocks5, aes(Lat, binomial)) +
  geom_rug() +
  geom_smooth(size = 0.5, color = "black", method = "glm", se = T, 
              method.args = list(family = "binomial")) +
  scale_x_continuous(name = "Latitude", expand = c(0,0), 
                     breaks = seq(64.15, 64.35, .05)) +
  scale_y_continuous(name = "Likelihood", 
                     expand = c(0, 0), 
                     limits = c(0, 1), 
                     breaks = seq(0, 1, .2)) +
  theme_classic() +
  theme (axis.title.y = element_text(size = 14, margin = margin(t = 0, r = 10, b = 0, l = 0), colour = "black"),
         axis.title.x = element_text(size = 14, margin = margin(t = 10, r = 0, b = 0, l = 0), colour = "black"),
         text = element_text(family = "Times New Roman"),
         legend.position = c(0.75,0.75),
         legend.text = element_text(size = 12),
         legend.title = element_text(size = 12),
         legend.key.size = unit(.5, "cm"),
         plot.margin = unit(c(0.2, 0.1, 0.1, 0.1), "cm"),
         axis.ticks.y = element_line(size = 0.5),
         axis.ticks.x = element_line(size = 0.5),
         axis.ticks.length = unit(0.2,"cm"),
         axis.text.y = element_text(colour = "black", size = 14, angle = 0, vjust = 0.5, hjust = 1,
                                    margin = margin(t = 0, r = 5, b = 0, l = 0)),
         axis.text.x = element_text(colour = "black", size = 14, angle = 90, vjust = 0, hjust = 0.5,
                                    margin = margin(t = 5, r = 0, b = 0, l = 0)),
         axis.line = element_line(colour = "black", size = 0.5, lineend = "square"),
         panel.border = element_rect(color = "black", fill = "transparent", size = 0.5))
```


```{r}
dev.new(width = 2.75, height = 1.83, units = "in", noRStudioGD = TRUE)
last_plot()

ggsave(file = "glm.tiff", width = dev.size()[1], height = dev.size()[2], dpi = 300)
```


